\textbf{Autoreduction - an automated data reduction system for beamline instruments at the ISIS Facility}

### 2017-07-11 Documentation/paper/paper.tex
Added name and paragraph about reduction to paper

\title{\textbf{Autoreduction - an automated data reduction system for beamline instruments at the ISIS Facility}}
\author{J. R. Corderoy, L. Greenwood, M. Noble, D. Oram and A. J. Markvardsen (code contributors)\\
ISIS Facility, Rutherford Appleton Laboratory\\
Email: someone with an STFC email account}
\date{23/09/2015}

@@

A key step for users of large research facilities is the
processing of instrument data before use. Autoreduction is a software
system designed to automate this process, integrated with other
pre-existing facility systems and software packages, and provide
an interface for users to control it.

This article describes the system setup for Autoreduction at the ISIS Facility;
it motivation and design, its use and interface, and its current status in production at ISIS.

@@

Large scale research facilities, such as the ISIS Neutron and Muon
Facility, allow users in academic and industrial research
to perform experiments using instruments set up to probe the inner 
structure and dynamics of materials. 
Common to most instruments is the need to process -- 
`reduce' -- raw data collected from an instrument into a form
that a user can interprete or can do further analysis on. For example
by removing instrument specific artifacts from the measured data, 
transforming the data into different units and/or
into data formats readably by further analysis software.

The purpose of the autoreduction system is to provide a useful automation
step in this process. The system described in this works allows data coming 
off an instrument to be automatically reduced using a dedicated software package 
such as Mantid\cite{mantid}, while lending transparency to the process and
letting users control it through a straightforward web interface.

TO DO: some reference that automation of data reduction also persued at
other large scale facilities: Anders writing from NOBUGS conference: such as 
possibly mention Diamand has a batch script based system 
for macromolecular beamlines (talk by Dr Markus Gerstel, Diamond). Dr Martin Rehr 
mentioned in his talk on "Image data Management System" that auto processing was
used. Dr Sergey Stepanov talk "Integration of fast detectors into.... macromolecular
crystallography beamlines at the Advanced Photon Source" used auto processing - seemed
to be a theme for MX beamlines.

@@

The autoreduction system at ISIS was inspired by an autoreduction setup at the
the Spallation Neutron Source (SNS) at Oak Ridge National Laboratory \cite{autoreductionSNS}.

The autoreduction implementation at ISIS uses the same underlying 
technologies (choice of programming language and queue node software
for example) as those used for the implementation of the SNS autoreduction
service but differs in that its front end client is designed to serve
autoreduction users only and in satisfying requirements of the ISIS 
facility not orginally part of the SNS implementation.

@@

The autoreduction service exists to reduce beamline instrument data. An
instrument produces discrete tranches of data, referred to as `runs', in
service of experiments. This raw data typically needs to be processed --
`reduced' -- before it can be used by scientists, by running a script on
it. Although the script itself need not change often, it may have
configurable parameters -- variables -- which might require more frequent
updating.

The autoreduction system is intended to automate this process.
Instrument scientists provide it with reduction scripts - in a specified
location on a shared file archive - and sets of variables to use to
reduce runs. The system watches for the completion of runs, and runs the
appropriate script, with the correct variables, on them, supplying the
reduced data to a data store for users to retrieve.

@@

The system is intended to serve the instrument scientists (who manage beamline instruments) 
and users who use the instruments to run experiments. 
The key requirements from these users, which have been impleted to date, are:

\begin{enumerate}[noitemsep]
\item
  Automate the reduction of data from beam instruments, using scripts
  provided by instrument scientists
\item
  Show the status of reduction on runs, as well as the results and
  locations of reduced data, to users
\item
  Allow instrument scientists to set the parameters used for reduction,
  in advance of experiments, and optionally schedule these to be diffirent
  for particular experiment
\item
  Allow users and instrument scientists to see what script, with which
  parameters, was used to reduce a run
\item
  Allow users and instrument scientists to modify the parameters used
  for a run, and re-run reduction using new parameters or scripts
\end{enumerate}

Additionally, constraints are:

@@

The architecture is based on a queue node, which co-ordinates
communication between all components of the system. This maintains
separate message queues, which allow components to act on messages from
other components and update the system. These components are

@@

>>>>

\title{\textbf{Autoreduction - an automated data reduction system for beamline instruments at the ISIS Facility}}
\author{J. R. Corderoy, T. Gowland, L. Greenwood, M. Noble, D. Oram and A. J. Markvardsen \\
(code contributors)\\
ISIS Facility, Rutherford Appleton Laboratory\\
Email: someone with an STFC email account}
\date{23/09/2015}

@@

A key step for users of large research facilities is the
processing of instrument data before use. Autoreduction is a software
system designed to automate this process, integrate with other
pre-existing facility systems and software packages, and provide
an interface for users to control it.

This article describes the system setup for Autoreduction at the ISIS Facility;
its motivation and design, its use and interface and its current status in production at ISIS.

@@

Large scale research facilities, such as the ISIS Neutron and Muon
Facility, allow users in academic and industrial research
to perform experiments using instruments set up to probe the inner 
structure and dynamics of materials. 
Common to most instruments is the need to process -- 
`reduce' -- raw data collected from an instrument into a form
that users can interprete or do further analysis on. For example
by removing instrument specific artifacts from the measured data, 
transforming the data into different units and/or
into data formats readable by further analysis software.

The purpose of the autoreduction system is to provide a useful automation
step in this process. The system described in this paper allows data coming 
off an instrument to be automatically reduced using a dedicated software package 
such as Mantid\cite{mantid}, whilst also lending transparency to the process and
letting users control it through a straightforward web interface.

Automated data reduction is a concept which is actively persued at other large facilities.
At the NOBUGS 2016 Conference\cite{NOBUGS}, there were several talks given by delegates
from other large scale facilities which mentioned the automated reduction of data at a variety
of facilities.

Data Analysis Infrastructure for DLS MX/CX \cite{dls-talk} discusses a batch
script based system for macromolecular beamlines. It makes use of ActiveMQ \cite{activemq}
(which will be discussed later in this paper) and also has a front end status monitor to
show the health of the system during run-time. Integration of Fast Detectors 
into Beamline Control \cite{anl-talk} also mentioned the use of automatic data processing.
To keep up with the data rates from faster detectors, they are now automatically reducing data
immediately so that users can see the results as soon as the dataset is taken.

@@

The autoreduction system at ISIS was inspired by an autoreduction setup at the
the Spallation Neutron Source (SNS) at Oak Ridge National Laboratory \cite{autoreduceSNS}.

The autoreduction implementation at ISIS uses the same underlying 
technologies (choice of programming language and queue node software
for example) as those used for the implementation of the SNS autoreduction
service. However, the ISIS implementation differs in that its front end 
client is designed to serve autoreduction users only and was not orginally 
part of the SNS implementation.

@@

The autoreduction service exists to reduce beamline instrument data. An
instrument produces discrete tranches of data, referred to as `runs', in
service of experiments. This raw data typically needs to be processed --
`reduced' -- before it can be used by scientists. This is done by running a 
script on it. Although the script itself need not change often, it may have
configurable parameters -- variables -- which might require more frequent
updating.

The autoreduction system is intended to automate this process.
Instrument scientists provide the system with reduction scripts - in a specified
location on a shared file archive - along with sets of variables to use to
reduce runs. Firstly, the system watches  for the completion of runs. Once a new
completed run has been found, the system will then run the appropriate script
with the correct variables on the new run. Once the script has finished execution,
reduced data will be saved to a data store for users to retrieve.

@@

The system is intended to serve the instrument scientists (who manage beamline instruments) 
and users who use the instruments to run experiments. 
The key requirements from these users, which have been completed to date, are:

\begin{enumerate}[noitemsep]
\item
  Automate the reduction of data from beam instruments, using scripts
  provided by instrument scientists
\item
  Show the status of reduction on runs, as well as the results and
  locations of reduced data, to users
\item
  Allow instrument scientists to set the parameters used for reduction
  in advance of experiments and optionally schedule these to be diffirent
  for particular experiment
\item
  Allow users and instrument scientists to see what script and with which
  parameters was used to reduce a run
\item
  Allow users and instrument scientists to modify the parameters used
  for a run and re-run reduction using new parameters or scripts
\end{enumerate}

Additional constraints are:

@@

The architecture is based on a queue node which co-ordinates
communication between all components of the system. This maintains
separate message queues which allow components to act on messages from
other components and update the system. These components are:

@@

\bibitem{dls-talk}
	M. Gerstel, R. Gildea, K. Levik, A. Ashton, G. Winter.
	\emph{Data Analysis Infrastructure for DLS MX/CX},
	NOBUGS Conference 2016, Københavns Universitet,
	\url{https://indico.esss.lu.se/event/357/session/7/contribution/51/material/slides/0.pdf}

\bibitem{anl-talk}
	Dr. Sergey Stepanov.
	\emph{Integration of fast detectors into beamline controls at the GM/CA macromolecular crystallography beamlines at the Advanced Photon Source},
	NOBUGS Conference 2016, Københavns Universitet,
	\url{https://indico.esss.lu.se/event/357/session/6/contribution/17/material/slides/0.pdf}

\bibitem{NOBUGS}
	NOBUGS Conference 2016,
	\emph{NOBUGS},
	\url{https://indico.esss.lu.se/event/357/}.
